---
title: "ML Final Project"
author: "Ian Yang"
date: "January 14, 2018"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(knitr)
library(gbm)
```

## Project Introduction
####Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


####Data

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

####Output
The goal of this project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


##Importing, cleaning and slicing data
```{r}
trainingData = read.csv('pml-training.csv', header = T, na.strings = c("NA", "", "#DIV/0!"))
testingData = read.csv('pml-testing.csv', header = T, na.strings = c("NA", "", "#DIV/0!"))
train = trainingData[,colSums(is.na(trainingData)) == 0]
test = testingData[,colSums(is.na(testingData)) == 0]
train = train[,-c(1:7)]
test = test[,-c(1:7)]
```

##Split data into traning/testing sets
```{r}
inTrain = createDataPartition(train$classe, p=0.7, list = FALSE)
myTrain = train[inTrain,]
myTest = train[-inTrain,]
```

##Apply Desicion Tree Algorithm
```{r}
set.seed(1234)
modFit1 = rpart(myTrain$classe ~ ., data = myTrain, method = "class")
fancyRpartPlot(modFit1)

prediction1 = predict(modFit1, myTest, type = "class")
cmtree = confusionMatrix(prediction1, myTest$classe)
cmtree
plot(cmtree$table, col = cmtree$byClass, main = paste("Decision Tree Confusion Matrix: Accuracy=", round(cmtree$overall['Accuracy'],4)))
```
##Apply Random Forests Algorithm
```{r}
modFit2 = randomForest(classe ~.,data = myTrain)
prediction2 = predict(modFit2, myTest, type = 'class')
cmrf = confusionMatrix(prediction2, myTest$classe)
cmrf
plot(modFit2)
```

##Applied Generalized Boosted Regression
```{r}
fitControl = trainControl(method = "repeatedcv", number = 5, repeats = 1)
gbmFit1 = train(classe ~., data = myTrain, method ="gbm", trControl = fitControl, verbose = FALSE)

gbmPredTest = predict(gbmFit1, newdata = myTest)
gbmAccuracyTest = confusionMatrix(gbmPredTest, myTest$classe)
gbmAccuracyTest
plot(gbmFit1)
```
Comparing the three ML algorithms above, random forests gave an accuracy of 99.35%, which is the highest among three options. The expected out-of-sample error is 100-99.35 = 0.65%
#Predicting outcome variable on the Test Data
```{r}
prediction3 = predict(modFit2, test, type = 'class')
prediction3
```





























